\documentclass[12pt]{report}
\usepackage{fullpage}
\title{Wormwood - A Text-Adventure Game System}
\author{Allan Lavell \\ Alex Moriarty \\ Jacob Godin \\ Nick Pinto}

\begin{document}
\maketitle
\setcounter{secnumdepth}{0}
\section{Introduction}
It's 1987: graphics haven't even been invented yet. You spend your nights sneaking into the local university computer lab to play text-adventure games on the latest hardware. 

Now, it's 2009: You can play text adventures on your brand-new quad-core gaming PC. This is where Wormwood steps in.

To say that our work on Wormwood has had any one specific "goal" would be
a bit of an over-simplification. Wormwood's overall goal is to make it easy 
for anyone to create their own text adventures, without putting any limits on
how they use the engine. In the context of our work on Wormwood over the time
period of the project, the main goal has been really "to get the basics working".
This goal has largely been achieved, and in a way that should make it possible
and maybe even easy to build upon in the future.

\section{Project Management}
Allan is the original creator of Wormwood. It was his idea to create the
engine, and he took on the role of project leader. Jacob is the most familiar
with Git, which is very useful source-control software, and he set up the
repository that we use for wormwood. Allan and Jacob then acted as gatekeepers
for the repository, and Nick and Alex sent their code to be checked and pushed
to the repo if it was okay to include. 

\section{Structure of the Source Code}
The code is divided into many separate 
classes, which are organized into packages. Here is a rundown of the packages:

\section{Packages}
\begin{itemize}
\item cmd: Contains all of the game commands. This directory is read by the parser to check to see if a particular command exists.
\item core: Contains the core components of the Wormwood, including the Parser and Output handler.
\item iface: Contains Interfaces to guide the creation of specific types of classes. Currently only contains the Command interface, which all commands must implement.
\item obj: Contains basic building block classes of the system, such as Room, Player, Item, and so forth.
\item game: Contains classes for building a game using Wormwood. Includes debugging features. 
\item doc: Contains documentation for the source code.
\end{itemize}

\section{Descriptions of Classes}
\subsection{Game}
The Game class has two important methods: prompt() and executeCommand(). prompt() gets the player's raw input from the terminal, and passes it over the Parser. The Parser returns the raw input as a Command object. executeCommand() takes this command object, and executes it by calling its exec method.

\subsection{Parser}
The Parser takes a raw String entered at the prompt. The general form for a command in Wormwood is \textit{cmd\_name arguments}. For example, the user might type "move e". The Parser splits this input up into the command and the arguments.  It then uses Java's generic class constructors to dynamically create the Command class, based on whether or not it can find the corresponding cmd\_name.java file in the cmd package. This simplifies things a lot, because it means that new commands can simply be added to the cmd directory without modifying the Parser to contain a list of available commands. It returns a generic Java Object containing the command. 

\subsection{Output}
Output encapsulates printing to the screen. At this point, Output.println() simply calls System.out.println(), but it allows for greater flexibility in the future. If we were to decide to make a more complicated output system that did things like make sure words didn't get cut off when they reach the edge of the screen, but instead were pushed down to the next line, we would only have to change the code in Output.


\subsection{Command}
Command isn't exactly a class - it's an Interface. All commands written for Wormwood should implement the Command interface. It enforces a few basic methods that all commands must have to work properly: exec, construct, and toString. The details of the implementation are left up to the class writer, but this helps give the programmer an idea of what is required. 

\subsection{Entity}
The base class for most of the objects in the game, Entity defines a few key methods and instance variables. Each entity has a name, a description (for when the entity is examined), a room description (what gets printed when the room is described to the player, i.e. "There's a chest in the corner of the room"), and an ArrayList of identifiers, which are names that can be used to identify the entity to commands. For example, an NPC (which extends the entity class) might have the room description "You see a man standing at a table in front of you." and the identifier "man". Then, when the player types "examine man", wormwood will reply with the man's description (e.g, "He's got a mean face, and blue jeans.").

\subsection{Player}
The Player class is an important class - it represents the player in the 
gameworld. It knows where the player currently is in the gameworld, and
also attributes of the player. It contains the player's inventory, as well
as their name, age, and gender. Basically the Player class keeps all the
information that will be with the character wherever he/she goes. Think of
it this way: the player will always be in a certain Room, so which room the
player is in will always be known by the player class. The player will always
have the same name, so that will also always be known by the Player class.

The Player class gets passed to the exec method of Commands by default,
because it contains so much pertinent information. 

\subsection{Room}
The basic building block in a gameworld created using Wormwood is a Room. Rooms inherit from the Entity class, mainly for the name and description variables and methods. Rooms are linked by Exits contained in the Grid (both of which are discussed elsewhere in this document). Rooms can contain items and NPC's, whose descriptions are given to the player along with the general description of the room.

\subsection{Exit}
Exits link rooms, and are member objects of the Grid (discussed below). Exits 
extend Entity, because they have names and descriptions that are displayed
to the player. They also have two significant data members: passable and 
locked. If an Exit is passable, that means the player can use the Exit
to go to wherever it points. If it is not passable, that means that the
player can see and examine the Exit, but cannot go through. An example of 
an object that is not passable is a closed door. Using the open command, the
player can make the Exit passable, if it isn't locked. If it is locked,
it has to be unlocked before it can become passable.

\subsection{Grid}
The Grid is an attempt to link the whole gameworld together. It consists of a 
two-element Array of points where Rooms can be stored, and an ArrayList of 
Exits that link those Rooms together. The ultimate goal of the Grid is to make
it easy to visualize where Rooms are in relation to each other in the 
gameworld. In the future, a GUI gamebuilder program could make use of the Grid
to display to the game maker how everything interconnects. 

Although not currently implemented, the idea is that there will be many
Grids in a game, each containing their own "level" or "area". The programmer
will then be able to link these Grids wherever and however they want to.
Grids are a level of organization above the basic building block of the Room.
They should help in cases like the following: the programmer is modeling a
house with wormwood. They want to have a main floor, a basement, and a top
floor. The main floor will have its own Grid, and be connected to the top 
floor, which is also on its own Grid. The basement Grid will be connected
to the main floor. This allows the programmer to break things up into
manageable sections.

\subsection{Item}
An Item is something that can be picked up by the player. It is then a part of
the player's inventory, and can be used, or dropped, at will. "Using" an item
merely invokes whatever command is associated with that particular item. For
example, if you had an item pencil, then you would make a command "Write" that
is associated with the pencil. Then, whenever the player "Uses" the pencil,
"Write" is invoked. Some items have a limited number of uses.

Currently commands mixed with items is not completely fleshed out. Because 
the commands that are associated with items right now are simply normal 
commands in the cmd package, they can be used independtly of the items they
are associated with, because the Parser automatically checks for everything
in the cmd package. This must be changed. A solution would be to have an
itemcmd package. This is most likely what will be implemented.

\subsection{NPC}
NPC's (Non-Player Characters) are right now basically just a direct extension of the Entity class.
They don't have any special features implemented for them yet, but ultimately
they will be the base class for characters in Wormwood. The player will be
able to interact with them, potentially doing such things as speaking to them,
trading/giving/receiving items with/from them, attacking them, being attacked
by them, etc.

\subsection{Create}
Currently, the way to create a game is through the Create class. In the class'
init() method, Grids, Entities, Rooms, etc., can all be defined by creating
Java objects from their classes. This leads to a lot of manual labour, and
will most definitely change. We are only now getting to the point where
we have enough of the framework done to really define how games will be 
created using said framework. Some possibilities include: a GUI
gamebuilder that automagically handles the creation of game entities in 
a neatly presented way; or writing a wormwood scripting language that simplifies
the syntax for creating things in wormwood. 

It is unfortunate that we haven't had more time to flesh this part of the 
system out, but as it stands, Wormwood works and it's manageable to create
a game. We decided to make sure we had enough of the framework done so 
that we could actually *make* games before we prettied everything up. 
Optimization on an unstable base is not a good strategy.



\end{document}
